UKB Simulation Analyses

### 2023-12-12 eda/ukb_curation-eda.Rmd
updating to make sure that ppts aren't resampled

source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source("/Users/megardn/Desktop/BGD_Repos/combat_biovar/R_scripts/gamlss_helper_funs.R")
```

@@

```{r}
cn_data <- fread(file="data/lifespan_dataset_dkatlas_CNonly.csv")
#table(cn_data$study)
names(cn_data)
```

@@

```{r}
col_list <- colnames(ukb_df)
pheno_list <- col_list[grepl("^lh_|^rh_|^GMV$|^WMV$|^sGMV$|^Ventricles$", col_list)]
length(pheno_list) #total number of models to run
saveRDS(pheno_list, file="R_scripts/pheno_list.rds")

#split by measure type (Vol, SA, CT)
vol_list_global <- c("GMV", "sGMV", "WMV", "Ventricles")
writeLines(vol_list_global, con = "pheno_lists/Vol_list_global.txt")
saveRDS(vol_list_global, file="R_scripts/vol_list_global.rds")

vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]
writeLines(vol_list_regions, con = "pheno_lists/Vol_list_regions.txt")
saveRDS(vol_list_regions, file="R_scripts/Vol_list_regions.rds")

sa_list <- col_list[grepl("^lh_SA|^rh_SA", col_list)]
writeLines(sa_list, con = "pheno_lists/SA_list.txt")
saveRDS(sa_list, file="R_scripts/SA_list.rds")

ct_list <- col_list[grepl("^lh_CT|^rh_CT", col_list)]
writeLines(ct_list, con = "pheno_lists/CT_list.txt")
saveRDS(ct_list, file="R_scripts/CT_list.rds")
```

@@

#write data
write_csv(ukb_df_totals, file="data/ukb_CN_data.csv")
```

@@

```{r}
ukb_df_age.filt <- ukb_df_totals %>%
  dplyr::filter(age_yrs >= 50 & age_yrs <= 80)
#write data
fwrite(ukb_df_age.filt, file="data/ukb_CN_data_agefilt.csv")
```

@@

#write out
write_csv(ukb_df_sites, file="data/ukb_to_model/ukb_CN_data_simsites.csv")
```

@@

## Simulate Sites - Permutations

# Create a list to store data frames
df_list <- list()

for(i in 1:5) {
  new_col_name <- paste0("SimSite", i)
  # Separate samples for males and females
  sampled_males <- sample(sim.site.list, size = n_male, replace = TRUE, prob = male_prob)
  sampled_females <- sample(sim.site.list, size = n_female, replace = TRUE, prob = female_prob)

  new_df <- ukb_df_age.filt %>%
    mutate(!!new_col_name := as.factor(ifelse(sex == "Male", sampled_males, sampled_females)))%>%
    dplyr::select(!!new_col_name)

  # Store the data frame in the list
  df_list[[i]] <- new_df
}

# Combine data frames into a single data frame
ukb_df_sites_perm <- bind_cols(ukb_df_age.filt, !!!df_list)

# View the final dataframe
#names(ukb_df_sites_perm)
```

```{r}
sim.perm.long <- ukb_df_sites_perm %>%
  dplyr::select(sex, age_days, contains("SimSite")) %>%
  pivot_longer(!c(sex, age_days), names_to = "permutation", names_prefix = "SimSite", values_to = "site")

prop <- sim.perm.long %>%
  group_by(permutation, site, sex) %>%
  summarize(Count = n(), avg_age = mean(age_days)) %>%
  group_by(permutation, site) %>%
  mutate(Proportion = Count / sum(Count))

summary.prop.male <- prop %>%
  dplyr::filter(sex=="Male") %>%
  group_by(site) %>%
  summarize(avg_prop = mean(Proportion), min_prop = min(Proportion), max_prop = max(Proportion))

summary.prop.male <- prop %>%
  group_by(site, sex) %>%
  summarize(avg_age_days = mean(avg_age), min_age = min(avg_age), max_age = max(avg_age))
```

```{r}
#write out
write_csv(ukb_df_sites_perm, file="data/ukb_CN_data_simsites_perm.csv")
```

>>>>

base_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/"

source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source(paste0(base_path, "cubic_scripts/R_scripts/gamlss_helper_funs.R"))
```

@@

```{r}
cn_data <- fread(file= paste0(base_path, "data/lifespan_dataset_dkatlas_CNonly.csv"))
#table(cn_data$study)
names(cn_data)
```

@@

```{r}
col_list <- colnames(ukb_df)
pheno_list <- col_list[grepl("^lh_|^rh_|^GMV$|^WMV$|^sGMV$|^Ventricles$", col_list)]
length(pheno_list) #total number of models to run
saveRDS(pheno_list, file=paste0(base_path, "cubic_scripts/R_scripts/pheno_list.rds"))

#split by measure type (Vol, SA, CT)
vol_list_global <- c("GMV", "sGMV", "WMV", "Ventricles")
writeLines(vol_list_global, con = paste0(base_path,"pheno_lists/Vol_list_global.txt"))
saveRDS(vol_list_global, file=paste0(base_path,"cubic_scripts/R_scripts/vol_list_global.rds"))

vol_list_regions <- col_list[grepl("^lh_Vol|^rh_Vol", col_list)]
writeLines(vol_list_regions, con = paste0(base_path,"pheno_lists/Vol_list_regions.txt"))
saveRDS(vol_list_regions, file=paste0(base_path, "cubic_scripts/R_scripts/Vol_list_regions.rds"))

sa_list <- col_list[grepl("^lh_SA|^rh_SA", col_list)]
writeLines(sa_list, con = paste0(base_path,"pheno_lists/SA_list.txt"))
saveRDS(sa_list, file=paste0(base_path,"cubic_scripts/R_scripts/SA_list.rds"))

ct_list <- col_list[grepl("^lh_CT|^rh_CT", col_list)]
writeLines(ct_list, con = paste0(base_path,"pheno_lists/CT_list.txt"))
saveRDS(ct_list, file=paste0(base_path,"cubic_scripts/R_scripts/CT_list.rds"))
```

@@

#write data
write_csv(ukb_df_totals, file=paste0(base_path,"data/ukb_CN_data.csv"))
```

@@

```{r}
ukb_df_age.filt <- ukb_df_totals %>%
  dplyr::filter(age_yrs >= 50 & age_yrs <= 80)
#write data
fwrite(ukb_df_age.filt, file=paste0(base_path,"data/ukb_CN_data_agefilt.csv"))
```

@@

#write out
write_csv(ukb_df_sites, file=paste0(base_path, "data/ukb_to_model/ukb_CN_data_simsites.csv"))
```

@@

### 2023-12-12 ukb_sim.site_analyses.Rmd
updating to make sure that ppts aren't resampled

```{r}
#load data csvs
ukb_df_cf.gam <- fread(file="data/ukb_to_model/cf.gam_data.csv")
ukb_df_cf.gamlss <- fread(file="data/ukb_to_model/cf.gamlss_data.csv")
ukb_df_cf.lm <- fread(file="data/ukb_to_model/cf.lm_data.csv")
```

>>>>

```{r}
#load data csvs
ukb_df_cf.gam <- fread(file="data/ukb_to_model/cf.gam_data.csv")
ukb_df_cf.gamlss <- fread(file="data/ukb_to_model/cf.gamlss_data.csv")
# length(ukb_df_cf.gamlss$participant) == n_unique(ukb_df_cf.gamlss$participant)
ukb_df_cf.lm <- fread(file="data/ukb_to_model/cf.lm_data.csv")
```
